Synthesis of Linear Functional Programs

### 2021-07-23 article/article.tex
Timeout on infinite evaluation, 5 hours lazy bug, $!, and reverse

\mypara{Optimizations} To speed up the process and get a cleaner and sometimes more correct
output, we add a rule that lets us ``skip some rules'' if left focused on a
$\bang$-ed proposition, and the goal is $\bang$-ed:
\begin{mathpar}
    \infer*[right=($\bang\Downarrow$L)]
    { \Theta/\Theta';\Rho;\Gamma;\Delta; A\Downarrow\ \vdash M : C}
    {\Theta/\Theta';\Rho;\Gamma;\Delta; \bang A\Downarrow\ \vdash M : \bang C}
\end{mathpar}

@@

\mypara{Memoization} Some proofs attempt to synthetise the same
sub-goal with the same premises more than once. Since these proof
attempts might be expensive performance-wise, and we can rely on the
program's determinism (equal calls to synthetise result in equal
outcomes), we added memoization to our synthesizer -- an optimization
technique that stores function calls and returns cached values for
equal inputs.
%
For our synthesizer, two calls are equivalent when
contexts $\Theta, \Rho, \Gamma, \Delta$, $\Omega$, (or alternatively
the left focus), and the goal, are equivalent. Because comparing all
of these is slow, our implementation hashes the relevant values via a hashing
library (\emph{Hashable}~\cite{https://hackage.haskell.org/package/hashable}) and
uses the result as the key.
%
Furthermore, a branch we want to memoize
might fail, not providing any result -- we also want to record
synthesis failure, as it happens more often than successful
results. We capture failure of a LogicT computation using the $\mathit{ifte}$
function and represent failure or success information with the \texttt{Maybe}
type. We ran benchmarks to measure the impact of memoization on 
performance. It's quite noticeable in synthesis with multiple
functions in the unrestricted context, or more complex
specifications. For example, if we add a new function \emph{read} to
the unrestricted context in the \emph{array} problem seen in
Section~\ref{sec:overview}, synthesis with memoization will take
around $5s$, while synthesis without memoization doesn't terminate
even after $15m$ running.
%
The implementation could be improved by better combining the
memoization with the LogicT monad transformer (remembering decisions across
backtracking branches) and by finetuning key-equality by ignoring context
components that aren't relevant to the synthesis call, to name some
examples. We leave such optimizations to future work.

>>>>

\mypara{Optimizations} To speed up the process and get a cleaner and sometimes more correct
output, we add a rule that lets us ``skip some rules'' if left focused on a
$\bang$-ed proposition, and the goal is $\bang$-ed:
\begin{mathpar}
    \infer*[right=($\bang\Downarrow$L)]
    { \Theta/\Theta';\Rho;\Gamma;\Delta; x{:}A\Downarrow\ \vdash M : C}
    {\Theta/\Theta';\Rho;\Gamma;\Delta; x{:}\bang A\Downarrow\ \vdash M : \bang C}
\end{mathpar}

@@

\mypara{Memoization} Some proofs attempt to synthetise the same
sub-goal with the same premises more than once. Since these proof
attempts might be expensive performance-wise, and we can rely on the
program's determinism (equal calls to synthetise result in equal
outcomes), we added memoization to our synthesizer -- an optimization
technique that stores function calls and returns cached values for
equal inputs.
%
For our synthesizer, two calls are equivalent when
contexts $\Theta, \Rho, \Gamma, \Delta$, $\Omega$, (or alternatively
the left focus), and the goal, are equivalent. Because comparing all
of these is slow, our implementation hashes the relevant values via a hashing
library (\emph{Hashable}~\cite{https://hackage.haskell.org/package/hashable}) and
uses the result as the key.
%
Furthermore, a branch we want to memoize
might fail, not providing any result -- we also want to record
synthesis failure, as it happens more often than successful
results. We capture failure of a LogicT computation using the $\mathit{ifte}$
function and represent failure or success information with the \texttt{Maybe}
type. We ran benchmarks to measure the impact of memoization on 
performance. It's quite noticeable in synthesis with multiple
functions in the unrestricted context, or more complex
specifications. For example, if we add a new function \emph{read} to
the unrestricted context in the \emph{array} problem seen in
Section~\ref{sec:overview}, synthesis with memoization will take
around $9s$, while synthesis without memoization doesn't terminate in less than
$15m$ running.
%
The implementation could be improved by better combining the
memoization with the LogicT monad transformer (remembering decisions across
backtracking branches) and by finetuning key-equality by ignoring context
components that aren't relevant to the synthesis call, to name some
examples. We leave such optimizations to future work.
